---
title: "Do Movie Genres Impact Ratings of Summer Movies?"
subtitle: "An Analysis Using IMDb Data"
author: "Jenna Evans"
institute: "Houghton University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(broom)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
tuesdata <- tidytuesdayR::tt_load('2024-07-30')

summer_movie_genres <- tuesdata$summer_movie_genres
summer_movies <- tuesdata$summer_movies

```

```{r}
style_xaringan(
  title_slide_background_image = "img/confetti.jpg"
)
```


--- 

class: center, middle

## Summer-Themed Movie Ratings on IMDb 

- Focused on Movies that were released during the summer season 
- On movies With "Summer" in the Title
- With the key question of how different geners of these movies influence IMDb ratings?


---


# Dataset Overview 

- Source: TidyTuesday (2024-07-30) 
- The dataset I used: `summer_movies`
- With the main Variables of: 
- `tconst`, `primary_title`, `average_rating`, `runtime_minutes`, `start_year`, `genres`


---

# Data Preparation

```{r preparation, echo = FALSE}
movies <- tuesdata$summer_movies %>%
  separate_rows(genres, sep = ",") %>%
  mutate(runtime_minutes = as.numeric(runtime_minutes)) %>%
  drop_na(average_rating, genres, runtime_minutes)
```

- Multiple genres split
- NAs dropped
- Cleaned and ready for analysis



--- 

# Summary Statistics by Genre

```{r summary-stats, echo = FALSE}

genre_summary <- movies %>%
  group_by(genres) %>%
  summarise(
    count = n(),
    mean_rating = mean(average_rating, na.rm = TRUE),
    median_rating = median(average_rating, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_rating))
genre_summary
```


---

class: middle, center

# Ratings by Genre (Boxplot) 

```{r boxplot-genre, fig.width=8, fig.height=5, echo = FALSE}
ggplot(movies, aes(x = fct_reorder(genres, average_rating), y = average_rating)) +
  geom_boxplot() +
  coord_flip() +
  theme_minimal() +
  labs(title = "IMDb Ratings by Genre", x = "Genre", y = "Average Rating")
```


---

# Ratings Over Time

```{r trend-year, fig.width=8, fig.height=5, echo = FALSE}
ggplot(movies, aes(x = year, y = average_rating)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", color = "blue") +
  theme_minimal() +
  labs(title = "Ratings Over Time", x = "Year", y = "Average Rating")
```


---

# Year vs. Rating: Regression

```{r regression-year, echo = FALSE}
model_year <- lm(average_rating ~ year, data = movies)
tidy(model_year)
```

- Slight negative trend
- Ratings may be declining slightly over time


---

# Key Insights

- Drama and Documentary genres rated highest, the highest rating though an outlier is a Talk-show film 
- Comedy and Horror show wide variability
- More recent movies trend slightly lower in rating
- Genre has a visible influence on IMDb scores


---

